An Experimental Evaluation of Hybrid Vector Search (EA\&B)

### 2025-04-30 VLDB2026/main.tex

	For third-party libraries and databases, Figure~\ref{fig:exp_8_2} shows that Faiss performs well in range filtering and even outperforms some specialized range query algorithms in certain cases. Faiss does not include optimizations specifically tailored for range filtering scenarios; its query mechanism is identical to that used in tag filtering scenarios. However, since attributes are pre-sorted in range filtering scenarios, the filtering overhead is minimal, leading to improved query performance. Similar to the selectivity observed in attribute filtering, Faiss performs better in small-range filtering scenarios because it searches over the filtered vectors. With a smaller range, fewer points remain after filtering, resulting in lower computational cost. In addition, the performance of Faiss is influenced by data dimensionality; it tends to work better on low-dimensional datasets.

@@

	It is important to note that this range query experiment is conducted under a fully static setting. Under such conditions, UNIFY, iRange, SeRF, and WinFilter require the dataset to be sorted by attribute prior to index construction. In contrast, DSG  support dynamic index construction with unordered attribute insertions, making them more suitable for streaming or incremental data scenarios. This flexibility highlights their superior extensibility in real-world applications.
	\subsubsection{Robustness.}

@@

	To evaluate the robustness of the RF-ANNS algorithm in different scenarios, we \textcolor{red}{add} an OOD dataset to the three standard datasets and \textcolor{red}{test} performance under large-range queries. As shown in Figure~\ref{fig:exp_8_3}, most methods \textcolor{red}{perform} worst on the Text2Image dataset. This is because OOD queries make the $k$-nearest neighbor distribution more sparse. As a result, the search needs to visit more nodes, which leads to an exponential growth in the search space and a sharp drop in efficiency.

>>>>

	For third-party \textcolor{red}{library} and databases, Figure~\ref{fig:exp_8_2} shows that Faiss performs well in range filtering and even outperforms some specialized range query algorithms in certain cases. Faiss does not include optimizations specifically tailored for range filtering scenarios; its query mechanism is identical to that used in \textcolor{red}{attribute} filtering scenarios. However, since attributes are pre-sorted in range filtering scenarios, the filtering overhead is minimal, leading to improved query performance. Similar to the selectivity observed in attribute filtering, Faiss performs better in small-range filtering scenarios because it searches over the filtered vectors. With a smaller range, fewer points remain after filtering, resulting in lower computational cost. In addition, the performance of Faiss is influenced by data dimensionality; it tends to work better on low-dimensional datasets.

@@

	It is important to note that this range query experiment is conducted under a fully static setting. Under such conditions, UNIFY, iRange, SeRF, and WinFilter require the dataset to be sorted by attribute prior to index construction. In contrast, DSG  supports dynamic index construction with unordered attribute insertions, making it more suitable for streaming or incremental data scenarios. This flexibility highlights its superior extensibility in real-world applications.
	\subsubsection{Robustness.}

@@

	To evaluate the robustness of the RF-ANNS \textcolor{red}{algorithms} in different scenarios, we \textcolor{red}{add} an OOD dataset Text2Image to the three standard datasets and \textcolor{red}{test} performance under large-range queries. As shown in Figure~\ref{fig:exp_8_3}, most methods \textcolor{red}{perform} worst on the Text2Image dataset. This is because OOD queries make the $k$-nearest neighbor distribution more sparse. As a result, the search needs to visit more nodes, which leads to an exponential growth in the search space and a sharp drop in efficiency.
